### 2018-11-02 pkg/model4you/inst/JORS/man_model4you.Rnw
comments Achim JORS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2374 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 33b6c34c30ee9921f076ef0bf033fcac905317a6

\section*{Abstract}
Typical models estimating treatment effects assume that the treatment effect is
the same for all individuals.  Model-based recursive partitioning allows to
relax this assumption and to estimate stratified treatment effects (model-based
trees) or even personalised treatment effects (model-based forests).  With
model-based trees one can compute treatment effects for different strata of
individuals. The strata are found in a data driven fashion and depend on
characteristics of the individuals.  Model-based random forests allow for a
similarity estimation between individuals in terms of model parameters (e.g.~
intercept and treatment effect). The similarity measure can then be used to
estimate personalised models.  The R package \emph{model4you} implements these
stratified and personalised models \new{in the setting with two randomly
assigned treatments} with a focus on ease of use and interpretability so that
clinicians and other users can take the model they usually use for the
estimation of the average treatment effect and with a few lines of code get a
visualisation that is easy to understand and interpret.

@@

A tree based on this model can be computed and visualised in only two lines of
code:
<<math_tree, fig.width = 13, fig.cap = "Personalised model tree for the Mathematics Exam datam.">>=
tr_math <- pmtree(bmod_math, control = ctree_control(maxdepth = 2))
plot(tr_math, terminal_panel = node_pmterminal(tr_math,
                                               plotfun = lm_plot))
@
\new{
We restrict the depth of the tree to two (\texttt{\Sexpr{'maxdepth = 2'}}) for
illustration purposes. If this setting is not used a more complex tree would be
estimated. Also the cutpoints and effect estimates are rounded in the plot.  To
view the tree in the R console, the command \texttt{\Sexpr{'print()'}} function can be
used.
}
<<math_printtree>>=
print(tr_math)
@
The tree (see Figure~\ref{fig:math_tree}) divides students based on the
percentage of successful online \texttt{\Sexpr{'tests'}}. These online tests were
conducted biweekly throughout the semester.  The largest difference between the
two exam groups is in the students who did very well in the online tests (more
than \texttt{\Sexpr{round(tr_math[[5]]$node$split$breaks, 1)}} percent correct). The
tree thus gives us much more information on the fairness of the exam than the
simple linear model, which is that it does not seem to be fair for students who
did very well throughout the semester (at this point we should state that the
students self selected into the two exam groups which might also be the reason
for differences in exam performance).
\new{Note that confidence intervals shown in the tree are not adjusted and
should hence be interpreted as a measure for variablity and not be used for
inference.}

>>>>

\section*{Abstract}
Typical models estimating treatment effects assume that the treatment effect is
the same for all individuals.  Model-based recursive partitioning allows to
relax this assumption and to estimate stratified treatment effects (model-based
trees) or even personalised treatment effects (model-based forests).  With
model-based trees one can compute treatment effects for different strata of
individuals. The strata are found in a data-driven fashion and depend on
characteristics of the individuals.  Model-based random forests allow for a
similarity estimation between individuals in terms of model parameters (e.g.~
intercept and treatment effect). The similarity measure can then be used to
estimate personalised models.  The R package \emph{model4you} implements these
stratified and personalised models \new{in the setting with two randomly
assigned treatments} with a focus on ease of use and interpretability so that
clinicians and other users can take the model they usually use for the
estimation of the average treatment effect and with a few lines of code get a
visualisation that is easy to understand and interpret.

@@

A tree based on this model can be computed and visualised in only two lines of
code:
<<math_tree, fig.width = 13, fig.cap = "Personalised model tree for the Mathematics Exam datam.">>=
tr_math <- pmtree(bmod_math, control = ctree_control(maxdepth = 2))
plot(tr_math, terminal_panel = node_pmterminal(tr_math,
                                               plotfun = lm_plot))
@
\new{
We restrict the depth of the tree to two (\texttt{\Sexpr{'maxdepth = 2'}}) for
illustration purposes. If this setting is not used a more complex tree would be
estimated. Also the cutpoints and effect estimates are rounded in the plot.  To
view the tree in the R console, the \texttt{\Sexpr{'print()'}} function can be
used.
}
<<math_printtree>>=
print(tr_math)
@
The tree (see Figure~\ref{fig:math_tree}) divides students based on the
percentage of successful online \texttt{\Sexpr{'tests'}}. These online tests were
conducted biweekly throughout the semester.  The largest difference between the
two exam groups is in the students who did very well in the online tests (more
than \texttt{\Sexpr{round(tr_math[[5]]$node$split$breaks, 1)}} percent correct). The
tree thus gives us much more information on the fairness of the exam than the
simple linear model, which is that it does not seem to be fair for students who
did very well throughout the semester (at this point we should state that the
students self selected into the two exam groups which might also be the reason
for differences in exam performance).
\new{Note that confidence intervals shown in the tree are not adjusted for
selection of the cutpoints in the tree and
should hence be interpreted as a measure for variablity and not be used for
inference.}
